$3$-D Weak Lensing Mass Map Reconstruction: Reducing the line of sight smearing

### 2020-12-09 doc/paper_ms_method_HSCY1/ms.tex
revise writing

\begin{abstract}
A new method is developed to reconstruct $3$-D density contrast maps from
photometric weak-lensing shear measurements. The $3$-D density contrast maps
are modeled as a summation of the NFW basis atoms, which have $2$-D multi-scale
NFW surface density profiles on the transverse plane and $1$-D Dirac delta
functions in the line of sight direction. With the prior assumption that the
density fields have a sparse spatial distribution, the density fields are
reconstructed using an oracle algorithm: adaptive lasso. The method is tested
with realistic simulations using the HSC-like shape estimation error and
photo-$z$ uncertainty.  Our findings are summarized as follows: 1) The lasso
solution suffers from a smear of structure in the line of sight direction even
in the absence of shape noise. In contrast, the adaptive lasso algorithm
efficiently removes the line of sight smear of structure.  2) The algorithm is
able to detect halo with minimal mass limits of $10^{14.0} M_{\odot}/h$,
$10^{14.7} M_{\odot}/h$, $10^{15.0} M_{\odot}/h$ for the low ($z<0.3$), median
($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts, respectively, with a
false detection of 0.022/deg$^2$.  3) The estimated redshift of the halos
detected from the reconstructed mass maps are lower than the true redshift by
about $0.03$ for halos at low redshifts ($z\leq 0.4$). The relative redshift
bias is below $0.5\%$ for halos at $0.4<z\leq 0.85$.
\end{abstract}

@@

The expected shear measurements ($\gamma$) on distant galaxies are related to
the foreground density contrast field ($\delta$) through a linear
transformation: \begin{equation} \label{eq-intro-delta2shear} \gamma=\mathbf{T}
\delta, \end{equation} where $\mathbf{T}$ is used to denote the linear
transformation operator, which includes not only the physical lensing effect but
also systematic effects from observations (e.g., pixelization and smoothing of
the shear field in the transverse plane, photo-$z$ uncertainty).

@@

The reconstruction of density map from shear measurements also receive
considerable interest as it reaches the nonlinear scales. The $2$-D density map
reconstruction which recovers an integration of projected mass along the line of
sight has been well studied within the community
\citep{massMap-KS1993,WL-massMap-Glimpse2D-Lanusse2016,sparseBaysianMassMap-Price2020}
and applied to large-scale surveys
\citep{HSC1-massMaps,massMapDES-Chang2018,DES-SV-massMap-sparsity}. However,
the reconstruction of $3$-D mass map is still a challenging task.

@@

Subsequently, in Section \ref{subsec_method_Systematics}, we discuss several
systematic effects from observations which include photo-$z$ uncertainty
(Section \ref{subsec_method_photoz}), smoothing (Section
\ref{subsec_method_smoothing}), masking(Section \ref{subsec_method_msknoise}),
and pixelization (Section \ref{subsec_method_pixel}).

@@

Combining equation (\ref{eq-delta2kappa}) with equation (\ref{eq-kappa2gamma}),
the expectation of lensing shear signal is
\begin{equation}\label{eq-delta2gammat}
\gamma_L(\vec{\theta},z_s) = \int_0^{z_s} dz_l K(z_l,z_s) \int d^2 \theta' \vec{D}(\vec{\theta}-\vec{\theta'}) \delta(\vec{\theta'},z_l).
\end{equation}

@@

We propose to use the multi-scale NFW atoms, denoted as
$\{\phi_1,...,\phi_N\}$, as the basis atoms of our dictionary.  On the
transverse plane, the NFW atoms follow surface density profiles of the NFW
halos \citep{haloModel-TJ2003-3pt} with comoving scale radii $r_s$ and
truncation radius $c r_s$, where $c$ is the concentration of the NFW
halos.  As the scales of halos are much smaller than the reachable redshift
resolution, we neglect the depth of halo on the line of sight direction and set
the NFW atoms' profiles in the line of sight direction to $1$-D Dirac delta
functions as suggested by \citep{LSS-massMap-Glimpse3D-Leonard2014}. The
multi-scale NFW atoms are defined as
\begin{equation}
\begin{split}
\phi_s(\vec{r}_{\theta},z) =&\frac{f}{2 \pi r_s^2 }
F(|\vec{r}_{\theta}|/r_s) \delta_D(z),\\
&  (s=1...N)
\end{split}
\end{equation}
where $\vec{r}_\theta$ is the comoving position in the transverse plane,
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh
\left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} +
\frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
$f=1/[\ln (1+c)-c/(1+c)]$. In this work, we fix $c=4$ for the NFW atoms in
different scale frames. Then, we write the basis atoms into the angular
separation coordinates:
\begin{equation}
\begin{split}
\phi_s(\vec{\theta},z) =&\frac{f \chi(z)^2}{2 \pi r_s^2 }
F(|\vec{\theta}|\chi(z)/r_s) \delta_D(z),\\
& (s=1...N)
\end{split}
\end{equation}
where $\chi(z)$ is the comoving distance at the redshift $z$.

@@

\begin{figure*}
\begin{minipage}[c]{1.0\columnwidth}
    \includezraphics{delta-3-7-pz-nn-lasso.pdf}
    \centering
    \small lasso
\end{minipage}
\begin{minipage}[c]{1.0\columnwidth}
    \includezraphics{delta-3-7-pz-nn-alasso.pdf}
    \centering
    \small adaptive lasso
\end{minipage}
\caption{The density map reconstructions with the lasso (left) and the adaptive
    lasso (right) algorithms. The vertical direction is the line of sight
    direction, and the lower boundaries and upper boundaries of the boxes are
    $z=0.01$ and $z=0.85$, respectively. Noises on galaxy shape measurements
    are neglected in the simulation.  The input halo mass is $M_{200}=10^{15}
    ~h^{-1}M_{\odot}$, and its redshift is $z=0.35$. The lasso reconstruction
    smears the density field along the line of sight direction.
    } \label{fig_lassoVsadaLasso}
\end{figure*}

@@

Figure \ref{fig_lassoVsadaLasso} shows the reconstructions of a single halo's
mass map with halo mass equals $M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift
$0.35$. The shear measurement error and photo-$z$ uncertainty are not included
in the simulation.  The left panel of Figure \ref{fig_lassoVsadaLasso} is the
reconstruction with the lasso algorithm, which shows a significant smear along
the line of sight.

>>>>

\begin{abstract}
A new method is developed to reconstruct three-dimensional density contrast
maps from photometric weak-lensing shear measurements. The three-dimensional
density contrast maps are modeled as a summation of the NFW basis atoms, which
have two-dimensional multi-scale NFW surface density profiles on the transverse
plane and one-dimensional Dirac delta functions in the line of sight direction.
With the prior assumption that the density fields have a sparse spatial
distribution, the density fields are reconstructed using an oracle algorithm:
adaptive lasso. The method is tested with realistic simulations using the
HSC-like shape estimation error and photometric redshift uncertainty.  Our
findings are summarized as follows: 1) The lasso solution suffers from a smear
of structure in the line of sight direction even in the absence of shape noise.
In contrast, the adaptive lasso algorithm efficiently removes the line of sight
smear of structure.  2) The algorithm is able to detect halo with minimal mass
limits of $10^{14.0} M_{\odot}/h$, $10^{14.7} M_{\odot}/h$, $10^{15.0}
M_{\odot}/h$ for the low ($z<0.3$), median ($0.3\leq z< 0.6$) and high
($0.6\leq z< 0.85$) redshifts, respectively, with a false detection of
0.022/deg$^2$.  3) The estimated redshift of the halos detected from the
reconstructed mass maps are lower than the true redshift by about $0.03$ for
halos at low redshifts ($z\leq 0.4$). The relative redshift bias is below
$0.5\%$ for halos at $0.4<z\leq 0.85$.
\end{abstract}

@@

The expected shear measurements ($\gamma$) on distant galaxies are related to
the foreground density contrast field ($\delta$) through a linear
transformation: \begin{equation} \label{eq-intro-delta2shear} \gamma=\mathbf{T}
\delta, \end{equation} where $\mathbf{T}$ is used to denote the linear
transformation operator, which includes not only the physical lensing effect
but also systematic effects from observations (e.g., pixelization and smoothing
of the shear field in the transverse plane, photometric redshift
uncertainty).

@@

The reconstruction of density map from shear measurements also receive
considerable interest as it reaches the nonlinear scales. The two-dimensional
($2$-D) density map reconstruction which recovers an integration of projected
mass along the line of sight has been well studied within the community
\citep{massMap-KS1993,WL-massMap-Glimpse2D-Lanusse2016,sparseBaysianMassMap-Price2020}
and applied to large-scale surveys
\citep{HSC1-massMaps,massMapDES-Chang2018,DES-SV-massMap-sparsity}. However,
the reconstruction of three-dimensional ($3$-D) mass map is still a challenging
task.

@@

Subsequently, in Section \ref{subsec_method_Systematics}, we discuss several
systematic effects from observations which include photometric redshift
(photo-$z$) uncertainty (Section \ref{subsec_method_photoz}), smoothing
(Section \ref{subsec_method_smoothing}), masking(Section
\ref{subsec_method_msknoise}), and pixelization (Section
\ref{subsec_method_pixel}).

@@

Combining eq. (\ref{eq-delta2kappa}) with eq. (\ref{eq-kappa2gamma}),
the expectation of lensing shear signal is
\begin{equation}\label{eq-delta2gammat}
\gamma_L(\vec{\theta},z_s) = \int_0^{z_s} dz_l K(z_l,z_s) \int d^2 \theta' \vec{D}(\vec{\theta}-\vec{\theta'}) \delta(\vec{\theta'},z_l).
\end{equation}

@@

We propose to use the multi-scale NFW atoms, denoted as
$\{\phi_1,...,\phi_N\}$, as the basis atoms of our dictionary.  On the
transverse plane, the NFW atoms follow surface density profiles of the NFW
halos \citep{haloModel-TJ2003-3pt} with comoving scale radii $r_s$ and
truncation radius $c r_s$, where $c$ is the concentration of the NFW halos.  As
the scales of halos are much smaller than the reachable redshift resolution, we
neglect the depth of halo on the line of sight direction and set the NFW atoms'
profiles in the line of sight direction to one-dimensional ($1$-D) Dirac delta
functions as suggested by \citep{LSS-massMap-Glimpse3D-Leonard2014}. The
multi-scale NFW atoms are defined as
\begin{equation}
\begin{split}
\phi_s(\vec{r}_{\theta},z) =&\frac{f}{2 \pi r_s^2 }
F(|\vec{r}_{\theta}|/r_s) \delta_D(z),\\
&  (s=1...N)
\end{split}
\end{equation}
where $\vec{r}_\theta$ is the comoving position in the transverse plane,
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh
\left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} +
\frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
$f=1/[\ln (1+c)-c/(1+c)]$. In this work, we fix $c=4$ for the NFW atoms in
different scale frames. Then, we write the basis atoms into the angular
separation coordinates:
\begin{equation}
\begin{split}
\phi_s(\vec{\theta},z) =&\frac{f \chi(z)^2}{2 \pi r_s^2 }
F(|\vec{\theta}|\chi(z)/r_s) \delta_D(z),\\
& (s=1...N)
\end{split}
\end{equation}
where $\chi(z)$ is the comoving distance at the redshift $z$.

@@

\begin{figure*}
\begin{minipage}[c]{1.0\columnwidth}
    \includezraphics{delta-3-7-pz-nn-lasso.pdf}
    \centering
    \small lasso
\end{minipage}
\begin{minipage}[c]{1.0\columnwidth}
    \includezraphics{delta-3-7-pz-nn-alasso.pdf}
    \centering
    \small adaptive lasso
\end{minipage}
\caption{The density map reconstructions with the lasso (left) and the adaptive
    lasso (right) algorithms. The vertical direction is the line of sight
    direction, and the lower boundaries and upper boundaries of the boxes are
    $z=0.01$ and $z=0.85$, respectively. Noises on galaxy shape measurements
    are neglected in the simulation.  The input halo's virial mass is
    $M_{200}=10^{15} ~h^{-1}M_{\odot}$, and its redshift is $z=0.35$. The lasso
    reconstruction smears the density field along the line of sight direction.
    } \label{fig_lassoVsadaLasso}
\end{figure*}

@@

Figure \ref{fig_lassoVsadaLasso} shows the reconstructions of a single halo's
mass map with virial mass equals $M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift
$0.35$. The shear measurement error and photo-$z$ uncertainty are not included
in the simulation.  The left panel of Figure \ref{fig_lassoVsadaLasso} is the
reconstruction with the lasso algorithm, which shows a significant smear along
the line of sight.
